\speculake: Hiding Malware in Speculative Execution

### 2018-05-08 paper/discussion.tex
Added spectre-NG mention and pass on future work

%\subsubsection{Full Whitebox Cryptography}
%To make it difficult to extract the key schedule used to decrypt the ciphertext,
%we can preload a large block of random data, and have multiple speculative entry
%points, each that choose a different index into the random data to be used as
%the key schedule, and then jumps to a common decryption routine. This way, the exact
%speculative entry point determines the key used for decryption. Because the
%exact speculative entry point is determined by the trigger program, an analyst
%cannot discover the decryption key directly from the payload program.
%
%An analyst could enumerate potential entry points, and test each one
%until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
%there are (at most) only about 1~million entry points, providing just 20~bits of
%security.
%
%To increase security, we could \emph{chain} entry points together. Rather than
%derive the index from the entry point, we have each potential entry point perform
%another indirect jump that the CPU cannot immediately resolve, forcing it to 
%speculate, inside the speculative world. The predicted
%target of that jump will also be determined by the trigger program. For
%instance, if the trigger program makes 30 training jumps, followed by 10
%additional jumps, the payload program will follow all 40 jumps, with the
%locations of the last 10 jumps used to determine the key index information.
%If each jump could land in up to 1024 possible locations, this would provide in
%total $1024^10 = 2^100$ possible keys, or 100~bits of security.
%
%A payload can still be crafted to use a benign program, given that it 
%implements a sufficient series of jumps. Even if a reverse engineer 
%gets to watch the program run they will only find out the plain text 
%and not the key schedule as any attempts to observe the decryption 
%process will interfere with the speculative branching. The reverse enginner 
%must either identify the exact process that was running concurently 
%which trained the branch predictor, or identify the correct series of 
%jumps taken speculatively to derive the key schedule. 
%
%We believe that the state explosion provided in both of these scenarios 
%presents new and difficult challenges for reverse engineers.

\subsubsection{Multiple Triggers}
To create further difficulties for a reverse engineer, or to further target
the execution environment it is possible
to have the payload program test a series of separate jump patterns. The decryption
process can then be done in stages where the result of a given stage determines 
the initial state of the next decryption stage that the payload program 
performs. Only when all triggers have been completed in the correct order will
the true plaintext be available.

@@

>>>>

This work demonstrates a general model for hiding execution in 
the speculative world and examines the implications and limitations on 
Intel processors. Given the wide-spread nature of the Spectre vulnerability 
and the ubiquity of cache side-channels we believe that this work can 
be directly extended to AMD, ARM, and other processors making use of
speculative branch prediction.

\subsubsection{Multiple Triggers}
To create further difficulties for an analyst, or to further target
the execution environment it is possible
to have the payload program test a series of separate jump patterns. The decryption
process can then be done in stages where the result of a given stage determines 
the initial state of the next stage that the payload program 
performs. Only when all triggers have been completed in the correct order will
the true plaintext be available.

@@

\smallskip

Alternatively multiple triggers could be used make individual state changes in
the payload program, and write in data from a local, or (as demonstrated in 
section~\ref{subsec:nested-spec}) remote processes. For example, a payload program
can check a series of three distinct jump patterns corresponding to three different 
triggers. If the first trigger is detected the payload shifts in a 1 to the state; if
the second is detected it shifts in a 0. If the third trigger is detected it 
discontinues listening to the three triggers and proceeds with other computation. 
This could be a direct form of \speculake using emulation and decryption 
or more traditional malicious behavior. Using this method a payload program can
be distributed without the encrypted data section all together, instead having 
the data written into the process at run-time without ever connecting out to any 
other resource.

\subsubsection{Benign Triggers}
We have demonstrated that a using OpenSSL as a benign trigger allows for 
\speculake payloads to be triggered remotely, and discussed the minimum 
requirements for identifying a benign trigger. However, we have not classified
typical system resources that might be used a benign triggers. This is a 
promising direction for future investigation as various system libraries present 
juicy targets for trigger design. For exmaple, if a sufficient series of jumps 
is found to exist in the TCP stack implemented in the unix kernel (i.e. in the TCP
Checksum library) an attacker could reasonably trigger a \speculake payload remotely,
from a spoofed IP address. This attack would rely on a payload defeating KASLR, 
however, this has been demonstrated to be feasible~\cite{gruss2017kaslr,evtyushkin2016jump}

\subsubsection{Alternative Applications}
We have proposed the \speculake model as a new malware threat, however, this model
of hidden computation applies more broadly. As of now there are no ways of 
debugging speculative computation meaning that the computation done there is effectively
isolated as with trusted execution environments. The security provided by the obfuscated 
key schedule makes \speculake useful in the domains of Digital Rights Management (DRM)
and whoitebox cryptography more generally. 
%% Approximate Computing
